---
title: "Interactive Maps of Tornado-Induced Property Damage"
author: "Jeremy Diaz"
date: "June 19, 2017"
output: html_document
---

### The Data

The data used for this visualization comes from the NOAA-provided Storm Events Database. The data is subsetted to only tornado events since Jan 1, 1997 with complete (no NA's) entries for property damage, beginning time & date, and beginning location.

### Map Details

The value which appears on the grey cluster points is indicative of how many tornado events are being clustered into that point. Each grey cluster point will decompose into individual tornado events upon sufficient zooming; this can be done by scrolling the mouse wheel, clicking the interactive buttons, or by clicking on the cluster point of interest. Tornadoes that caused extreme damage appear yellow; tornadoes that caused intermediate damage appear green (more towards extreme) and blue (more towards no damage); and, tornadoes that caused no damage appear dark purple.

If you hover your mouse over a tornado event point, a label will display the property damage caused by that event and the date when the damage occurred.

```{r interactive_map_plot_clusters, echo = FALSE, message = FALSE}


# Packages
library(leaflet)
library(viridis)
library(mapview)


# Import the data
tor_df <- read.csv("~/earth-analytics/tornadoesr/data/raw/leaflet_viz_data.csv")


# Make log-transformed property damage its own column
tor_df$LOG_DAM <- log(tor_df$DAMAGE_PROPERTY + 1,
                      base = exp(1))


# Get the date on each event (without time)
tor_df$BEGIN_DATE <- substring(tor_df$BEGIN_DATE_TIME, 1, 10)


# Get the color a palette
pal <- colorNumeric(palette = viridis(22),
                    domain = tor_df$LOG_DAM)


# No scientific "e+6" notation
options("scipen" = 100)


# Make the interactive map with clusters
leaflet(tor_df,
        width = "100%",
        padding = 3,
        options = leafletOptions(minZoom = 4)) %>%
  addTiles() %>%
  addProviderTiles(providers$CartoDB.DarkMatter) %>%
  addCircleMarkers(lng = tor_df$BEGIN_LON,
                   lat = tor_df$BEGIN_LAT, 
                   stroke = FALSE,
                   radius = 15,
                   fill = TRUE,
                   fillColor = ~pal(tor_df$LOG_DAM),
                   fillOpacity = 1,
                   labelOptions = labelOptions(textsize = "15px"),
                   label = ~paste(paste("Property Damages: ",
                                        paste0(paste0("$",
                                                      prettyNum(tor_df$DAMAGE_PROPERTY,
                                                                big.mark = ",",
                                                                scientific = FALSE)),
                                               ";")),
                                  paste("Date: ",
                                        tor_df$BEGIN_DATE),
                                  sep = "   "),
                   clusterOptions = markerClusterOptions(radius = 15,
                                                         iconCreateFunction = JS("function (cluster) {    
    var childCount = cluster.getChildCount();  
    if (childCount < 100) {  
      c = 'rgba(135, 135, 135, 1.0);'
    } else if (childCount < 1000) {  
      c = 'rgba(135, 135, 135, 1.0);'  
    } else { 
      c = 'rgba(135, 135, 135, 1.0);'  
    }    
    return new L.DivIcon({ html: '<div style=\"background-color:'+c+'\"><span>' + childCount + '</span></div>', className: 'marker-cluster', iconSize: new L.Point(40, 40) });

  }"))) %>%
  addLegend("bottomright",
            pal = pal,
            values = ~tor_df$LOG_DAM,
            opacity = 1,
            title = "Log-Transformed <br> Property Damage <br> (US dollars + 1)")


```

### Slightly Modified

The following is the same map as the above with two changes: 

1. There are no clusters 
2. Point sizes are scaled down for better large-scale viewing

```{r interactive_map_no_clusters, echo = FALSE}


# Make the interactive map with clusters
leaflet(tor_df,
        width = "100%",
        padding = 3,
        options = leafletOptions(minZoom = 4)) %>%
  addTiles() %>%
  addProviderTiles(providers$CartoDB.DarkMatter) %>%
  addCircleMarkers(lng = tor_df$BEGIN_LON,
                   lat = tor_df$BEGIN_LAT, 
                   stroke = FALSE,
                   radius = 3,
                   fill = TRUE,
                   fillColor = ~pal(tor_df$LOG_DAM),
                   fillOpacity = 0.95,
                   labelOptions = labelOptions(textsize = "15px"),
                   label = ~paste(paste("Property Damages: ",
                                        paste0(paste0("$",
                                                      prettyNum(tor_df$DAMAGE_PROPERTY,
                                                                big.mark = ",",
                                                                scientific = FALSE)),
                                               ";")),
                                  paste("Date: ",
                                        tor_df$BEGIN_DATE),
                                  sep = "   ")) %>%
  addLegend("bottomright",
            pal = pal,
            values = ~tor_df$LOG_DAM,
            opacity = 1,
            title = "Log-Transformed <br> Property Damage <br> (US dollars + 1)")


```